---
title: "Plotting and Summarizing Activity Data"
author: "Samyam"
date: "6/5/2020"
output: html_document
keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading dependency and loading the activity.csv file
Modifying the structure of date column in a
```{r}
library(tidyverse)
a<-read.csv("./activity.csv", sep=",", header = T)
a$date<-as.Date(a$date)
```
## What is mean total number of steps taken per day?

### Calculate the total number of steps taken per day
```{r}
b<-a%>%group_by(date)%>%filter(is.na(steps)!=T)%>%summarize(sum_steps_day=sum(steps,na.rm = T))
head(b,10)
```

### Make a histogram of the total number of steps taken each day
```{r}
ggplot(b, aes(x=date, y=sum_steps_day))+
  geom_histogram(stat="identity", fill="blue",col="red",binwidth=30, width=NULL)+
  scale_x_date(date_breaks = "5 days", date_labels =  "%d %b")+
  theme(axis.text.x=element_text(angle=60, hjust=1), panel.grid = element_blank(), panel.background = element_blank())+
  xlab("Day") + 
  ylab("Sum of steps per day") + 
  ggtitle("total number of steps taken per day")
```
### Calculate and report the mean and median of the total number of steps taken per day
```{r}
summary(b$sum_steps_day)
```
## What is the average daily activity pattern?

### Make a time series plot (i.e.type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
c<-a%>%group_by(interval)%>%filter(is.na(steps)!=T)%>%summarize(mean_steps_day=mean(steps,na.rm = T))

```

```{r}
plot(x=c$interval, y=c$mean_steps_day, type="l", lwd=2,col="red",ylab = "mean steps per day", xlab="interval of 5 minutes", main="mean steps per day per 5 minute interval")
```

### Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
e<-c[which.max(c$mean_steps_day),]
paste("The maxinum number of mean step is in the interval", e[1,1], "and the mean step during that interval is", e[1,2])
```

## Imputing missing values
### Calculate and report the total number of missing values in the dataset (i.e. the total number of rows)
```{r}
sum(is.na(a$steps))
```
### Choosing the mean of 5 minute interval and replacing them
### Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
newa<-a
for (i in 1:nrow(newa)) {
  if(is.na(newa$steps[i])) {
    meansperinterval <- round(c$mean_steps_day[which(c$interval == newa$interval[i])])
    newa$steps[i] <- meansperinterval
  }
}
```

### Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
The imputing increases the overall mean of the data set
It also places numerous bias
```{r}
sum(is.na(newa$steps))
```
```{r}
f<-newa%>%group_by(date)%>%summarize(sum_steps_day=sum(steps,na.rm = T))
ggplot(f, aes(x=date, y=sum_steps_day))+
  geom_histogram(stat="identity", fill="blue",col="red")+
  scale_x_date(date_breaks = "5 days", date_labels =  "%d %b")+
  theme(axis.text.x=element_text(angle=60, hjust=1), panel.grid = element_blank(), panel.background = element_blank())+
  xlab("Day") + 
  ylab("Sum of steps per day") + 
  ggtitle("total number of steps taken per day")
```
Reporting the mean of the new dataset
```{r}
summary(f$sum_steps_day)
```

## Are there differences in activity patterns between weekdays and weekends?
### Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
newa$day<-ifelse(weekdays(newa$date)=="Saturday"|weekdays(newa$date)=="Sunday", "Weekend", "Weekday")

```
```{r}
date<-aggregate(steps ~ interval+day, newa, mean)
head(date)
```

### Make a panel plot containing a time series plot (type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r}
ggplot(date, aes(x=interval, y=steps))+geom_line(size=0.75,aes(col=day))+facet_wrap(.~day)
```
